Texture Mapping 3D Models of Indoor Environments with Noisy Camera Poses

### 2012-06-25 3dimpvt/paper.tex
fixed first half, need more refs and waiting for second half

%%%%%%%%% ABSTRACT
\begin{abstract}
  Automated 3D modeling of building interiors is useful in
  applications such as virtual reality and environment
  mapping. Applying textures to these models is an important step in
  generating photorealistic visualizations of data gathered by
  modeling systems.  The localization of cameras in such systems often
  suffer from inaccuracies, resulting in visible discontinuties when
  different images are projected adjacently onto a plane for
  texturing. We propose two approaches for reducing discontinuities
  during texture mapping, one to robustly accomodate images of all
  orientations, and one to take advantage of optimal situations where
  images have uniform orientation. The effectiveness of our approaches
  will be demonstrated on two indoor datasets.
\end{abstract}

@@

The aim of this paper is to present a solution for texture mapping the
3D models generated by indoor modeling systems, with specific
attention given to a human-operated system with high camera pose
errors and great variance in camera locations.

The first step in automated 3D modeling is the physical scanning of
the environment's geometry. An indoor modeling system must be able to
recover camera poses within an environment while simulatenously
reconstructing the 3D structure of the environment itself. This is
known as the simultaneous localization and mapping (SLAM) problem, and
is generally solved by taking readings from laser range scanners,
cameras, and inertial measurement units (IMUs) at multiple locations
within the environment.

Mounting such devices on a human-carried platform provides unique
advantages over vehicular-based systems in terms of agility and
portability. Unfortunately, human-operated systems also result in much
larger localization error. As a result, common methods for texture
mapping generally produce poor results, as later shown in Sections
\ref{sec:simpleTextureMapping} and
\ref{sec:existingApproaches}. Before discussing how to overcome these
challenges, we first provide an overview of the backpack modeling
system from which our test data has been obtained.

Our backpack-mounted modeling system contains five 2D laser range
scanners, two cameras, and an orientation sensor. The laser scanners
are mounted orthogonally and have a 30-meter range and a 270$^{\circ}$
field of view. The two cameras are equipped with fisheye lenses,
reaching an approximately 180$^{\circ}$ field of view, and are mounted
with one facing left and the other facing right. These cameras take
images at the rate of 5 Hz. The orientation sensor provides
orientation parameters at a rate of 180 Hz.

@@

Using data gathered by the onboard sensors and multiple localization
and loop-closure algorithms, the backpack is then localized over its
data collection period, and a 3D point cloud of the surrounding
environment is constructed \cite{chen2010indoor, kua2012loopclosure,
  liu2010indoor}. Approximate normal vectors for each point in the
point cloud are then calculated by gathering neighboring points within
a small radius and processing them through principal component
analysis. These normal vectors allow for the classification and
grouping of adjacent points into structures such as walls, ceilings,
floors, and staircases. A RANSAC algorithm is then employed to fit
polygonal planes to these structured groupings of points, resulting in
a fully planar model \cite{sanchez2012point}. This model, consisting
of multiple 2D polygonal planes in 3D space, along with the set of
images captured by the backpack's cameras and their 3D poses, can be
considered the input to our texture mapping problem

The remainder of the paper is organized as follows. Section
\ref{sec:simpleTextureMapping} describes the general problem of
texture mapping and examines two simple mapping approaches and their
shortcomings.  Section \ref{sec:existingApproaches} demonstrates two
previous attempts at localization refinement, and demonstrates their
inadequacies for our datasets.  Section \ref{sec:proposedApproach}
presents our proposed approach to texture mapping, combining an
improved localization refinement process with two image selection
approaches. Section \ref{sec:resultsAndConclusions} contains results
and conclusions.

\section{Simple Texture Mapping}
\label{sec:simpleTextureMapping}
\begin{figure}
  \centering
  \includegraphics[height=2in]{Projection.pdf}
  \caption{Planes are specified in 3D space by four corners $C_1$ to
    $C_4$. Images are related to each plane through the camera
    matrices $P_{1..M}$. }
  \label{fig:projection}
\end{figure}

@@

The geometry of the texture mapping process for a plane is shown in
Figure \ref{fig:projection}.  As described earlier, we are provided
with a set of $M$ images with which we must texture our target
plane. Each image has a camera matrix $P_i$ for $i=1..M$, which
translates a 3D point in the world coordinate system to a 2D point or
pixel in image $i$'s coordinates. If the 3D world point is not
contained in the image, the 2D point will simply be outside of the
image boundaries. A camera matrix $P_i$ is composed of the camera's
intrinsic parameters, such as focal length and image center, as well
as extrinsic parameters which specify the rotation and translation of
the camera's position in 3D world coordinates at the time that image
$i$ is taken. These extrinsic parameters are determined by the
backpack hardware and localization algorithms mentioned earlier. A
point $X$ on the plane in 3D space can be related to its corresponding
pixel $x$ in image $i$ through the following equation:

\[
x=project(P_iX)
\]

where
\[X = \begin{pmatrix} x \\ y \\ z \end{pmatrix} \textrm{ and }
project(X) = \begin{pmatrix} x/z \\ y/z \end{pmatrix}
\]

For the sake of simplicity, we treat all planes as rectangles by
generating minimum bounding boxes for them. Since our final textures
are stored as standard rectangular images anyway, we can simply leave
the area between plane boundary and bounding box untextured, or crop
it out as needed. A plane to be textured is thus defined by a bounding
box with corners $C_i$ in world coordinates and a normal vector
indicating the front facing side of the plane. Our goal is to texture
this plane using images captured by the backpack system, while
eliminating any visual discontinuities or seams that would suggest
that the plane's texture is not composed of a single continuous image.

@@

In order to select an image for texturing tile $t$, we must first
gather a list of candidate images that contain all four of its
corners, which we can quickly check by projecting $t$ into each image
using the projection method above. Furthermore, each candidate image
must have been taken at a time when its camera had a clear
line-of-sight to $t$, which can be calculated using standard
ray-polygon intersection tests between the camera location, the center
of $t$, and other planes, all in world coordinates.

Once we have a list of candidate images for $t$, we must define a
scoring function in order to compare images and objectively select the
best one. Since camera localization errors compound over distance, we
wish to minimize the distance between cameras used for texturing and
our plane. Additionally, we desire images that are projected
perpendicularly onto the plane, maximizing the resolution and amount
of useful texture available in their projections.

These two criteria can be met by maximizing the function shown in
Figure \ref{fig:scoringFunction}.

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{scoringFunction.pdf}
  \caption{Images with minimal camera angle $\alpha$ and distance $d$
    are selected by maximizing the function}
  \begin{center}$\frac{1}{d} \cdot (-1 \cdot C) \cdot N$ \end{center}
  \label{fig:scoringFunction}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=3in]{wall1_naive.jpg}
  \caption{The result of direct texture mapping based on locally
    optimized textures.}
  \label{fig:directMapping}
\end{figure}

As Figure \ref{fig:directMapping} demonstrates, this approach leads to
the best texture for each tile independently, but overall results in
many image boundaries with abrupt discontinuities, due to significant
misalignment between images.

\subsection{Mapping with Caching}
\label{sec:mappingWithCaching}
Since discontinuities occur where adjacent tiles select different
images that do not align, it makes sense to take into account image
selections made by neighboring tiles while selecting the best image
for a given tile. By using the same image across tile boundaries, we
eliminate the discontinuity altogether. If this is not possible
because a tile is not visible in images chosen by its neighbors, using
similar images is likely to result in less noticeable discontinuities.

Similar to a caching mechanism, we select the best image for a tile
$t$ by searching through two subsets of images for a good candidate,
before searching through the entire set. The first subset of images is
those selected by adjacent tiles that have already been textured. We
must first check which images can map to $t$, and then of those, we
make a choice according to the same scoring function in
\ref{fig:scoringFunction}. Rather than blindly reusing this image, we
ensure it meets a threshold, which we set to $\alpha < 45^\circ$, to
be considered a good image.

If no satisfactory image is found, we then check our second subset of
images, which consists of images that were taken near the images in
the first subset, both spatially and temporally. These images are not
the same as the ones used for neighboring tiles, but they were taken
at a similar location and time, suggesting that their localization and
projection are very similar. Again, if no good image is found
according to the same threshold, we must then search the entire set of
candidate images.

\begin{figure}
  \centering
  \includegraphics[width=3in]{wall1_cache_full.jpg}
  \caption{The result of adding a caching system to locally optimized
    textures.}
  \label{fig:caching}
\end{figure}

The result of this caching approach is shown in Figure
\ref{fig:caching}. As compared to Figure \ref{fig:directMapping},
discontinuities have been reduced overall, but the amount of remaining
seams suggests that image selection alone cannot produce seamless
textures. Camera matrices, or the image projections themselves have to
be adjusted in order to reliably generate clean textures.

\section{Existing Approaches to Image-Aligned Texture Mapping}
\label{sec:existingApproaches}
In order to produce seamless texture mapping, either camera matrices
need to be refined such that their localization is pixel accurate,
resulting in a perfect mapping, or image stitching techniques need to
be applied to provide this illusion. Before examining these
approaches, we first obtain a set of images to work with.

Rather than perform camera or image adjustments across the many
thousands of images acquired in a typical data collection, we opt to
work with the more limited set of images corresponding to those chosen
by the direct mapping approach, without caching. This set of images
constitutes a good candidate set for generating a final seamless
texture since it meets three important criteria. First, each tile on our planeis covered by at least one image in this set; this ensures no
holes in our final texture. Second, since images are all selected
according to the same scoring function in Figure
\ref{fig:scoringFunction}, they are taken at as much of a head-on
angle as possible and should project onto the plane in similar
ways. Third, as a side result of the scoring function, selected images
are only good candidates for the tiles near their center of
projection. Thus, there should be plenty of overlap between selected
images, allowing for some degree of shifting without resulting in
holes, as well as area for blending between them. With this set of
images, we now review two existing approaches towards refining and
combining their projections.

\subsection{Image Mosaicing}
\label{sec:imageMosaicing}
\begin{figure}
  \centering
  \includegraphics[width=3in]{panoMy.jpg}
  \caption{Image mosaicing. }
  \label{fig:mosaic}
\end{figure}

@@

Figure \ref{fig:mosaic} shows the output of the AutoStitch software
package which does homography-based image mosaicing. This plane is
nearly a best-case scenerio with many features spread uniformly across
it. Even so, the mosaicing produces errors that causes straight lines
to appear as waves on the plane. This image was generated after
careful hand tuning as well. Many planes that had fewer features
simply failed outright. Thus, image mosaicing is not a robust enough
solution for reliably texture mapping our dataset.

\subsection{Image-Based 3D Localization Refinement}
\label{sec:imageBased3DRefinement}
\begin{figure}
  \centering
  \includegraphics[width=3in]{Graph_crop.pdf}
  \caption{The graph-based localization refinement algorithm from \cite{chen2010indoor}
    suffers from the problem of compounding errors. }
  \label{fig:graph}
\end{figure}

Another approach is to refine the camera matrices using image
correspondences to guide the process. Each image's camera matrix has 6
degrees of freedom that can be adjusted. Previous work on this problem
attempted to refine camera matrices by solving a non-linear
optimization problem \cite{liu2010indoor}. This process is specific to
the backpack system which generated our dataset, as it must be run
during backpack
localization\cite{liu2010indoor, chen2010indoor}. Unfortunately, this
approach suffers from a similar error propagation problem as shown in
Figure \ref{fig:graph}.

\begin{figure}
  \centering
  \includegraphics[width=3in]{pipeline.pdf}
  \caption{Our proposed method for seamless texture mapping.}
  \label{fig:pipeline}
\end{figure}
\section{Proposed Method for Seamless Texture Mapping}
\label{sec:proposedApproach}
In this section, we will describe our proposed method for seamless
texture mapping. Our approach contains 4 steps, as seen in
\ref{fig:pipeline}.  First, we obtain a set of images to work with, as
already described in Section \ref{sec:existingApproaches}. Next, we perform
image rotation and shifting in order to maximize SIFT feature matches
in overlapping areas between images. We then project these images onto
our plane, applying textures either with the tile caching method from
Section \ref{sec:mappingWithCaching}, or with the more specialized
method described in Section \ref{sec:seamMinimization}. We then finish
by applying linear alphar blending to smooth out any remaining seams.

\subsection{Image Projection and Rotation}
\label{sec:projectionAndRotation}
Our localization approach begins with the projection of all our images
onto separate copies of our plane, such that no projected data is
covered up and lost. This is done in the same way as the approaches in
Section \ref{sec:simpleTextureMapping}. We then perform rotations on
these projections, as adjacent images with different orientations will
result in strong discontinuities.

These rotations are accomplished by using Hough transforms, which
detect the presence and orientation of linear features in our
images. Rather than match the orientation of such features in each
image, we simply apply rotations such that the strongest near-vertical
features are made completely vertical. This is effective for indoor
models, since strong features in indoor scenes usually consist of
parallel vertical lines corresponding to doors, wall panels,
rectangular frames, etc. If features in the environment are not
vertical, or are not parallel to eachother, this step is skipped.

\begin{figure}
  \centering
  \includegraphics[width=3in]{matches.jpg}
  \caption{SIFT feature matches between overlapping images.}
  \label{fig:matches}
\end{figure}

\subsection{Robust SIFT Feature Matching using RANSAC}
\label{sec:robustSIFTFeatureMatching}
Our next step is to fix misalignment between overlapping images. We do
this by first searching for corresponding points between all pairs of
overlapping images using SIFT feature matching
\cite{lowe1999object}. An illustration of this is given in Figure
\ref{fig:matches}. The SIFT matches allow us to determine $dx$ and
$dy$ distances between each pair of features for two images on the
plane, determining where they should be projected relative to
eachother.

@@

% Draw least squares problem here.

@@

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and match each other
with far greater accuracy. Applying the tile caching method from
Section \ref{sec:mappingWithCaching} on these re-localized images
results in the significant improvements shown in Figure
\ref{fig:shifted}, as compared to Figure \ref{fig:caching}.

\begin{figure}
  \centering
  \includegraphics[width=3in]{wall1_cache_full_shifted.jpg}
  \caption{Localization refinement results in signifcantly fewer
    discontinuities in the final texture.}
  \label{fig:shifted}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=3in]{highCameraAngle.pdf}
  \caption{The image on the left projects head-on not only onto point A, but most visible areas of the wall that point A is on. The image's projection onto point B will be at a high angle, but it will still be much more usable than the projection onto point C.}
  \label{fig:highCameraAngle}
\end{figure}

As mentioned earlier, our backpack system has 2 cameras with
$180^\circ$ fisheye lenses facing to the right and left. Since the
backpack operator only takes care to fully scan each wall and not
necessarily the entirety of each ceiling and floor, cameras at higher
angles with respect to plane normal vectors must be used for texturing
large areas of floor and ceiling planes. These high camera angles
translate into images that span extremely large areas once
projected. Such image projections have good scores (from Figure
\ref{fig:scoringFunction}) for certain areas on the plane, but much
worse scores for other areas. This can be seen in Figure
\ref{fig:highCameraAngle}. The tiling approach used in Section
\ref{sec:simpleTextureMapping} was thus a good choice for texturing,
as it allowed us to only use the parts of image projections that were
good choices for their respective plane locations.

For wall planes however, we have images all taken from close distances
and more head-on angles, and thus much smaller near-rectangular
projections to work with. As a result, there is less deviation over
the score of each tile within an image, as well as over the average
scores of all images. This means that the scoring criteria from Figure
\ref{fig:scoringFunction} is less relevant to walls, which enjoy an
abundance of head-on images.

@@

\subsection{Occlusion Masking}
\label{sec:occlusionMasking}
To begin with, we need to ensure that our images contain only content
that should be mapped onto the target plane in question. The tiling
approach used previously only checks occlusion for each tile as it is
being textured. For our new approach, we prefer entire images, so we need
to perform occlusion checks over the entirety of each image to
determine available areas for texture mapping.

@@

\subsection{Image Selection}
\label{sec:imageSelection}
To determine the set of images that results in the cleanest texture,
we need a cost function to evaluate the visibiilty of seams between
images in our set. A straightforward cost function that accomplishes
this is the sum of squared pixel differences in overlapping regions
between all pairs of images after they have been aligned as described
in Section \ref{sec:projectionAndRotation}. Minimizing this cost
function encourages image boundaries to occur either in featureless
areas, such as bare walls, or in areas where images match extremely
well.

With the cost function defined, we must now select the set of images
for which the overall cost function is minimized. Since nearly all the
images for our optimal case are head on, the best strategy to minimize
seams is to choose as few images as possible while texturing a given
plane. Thus,to cover the entirety of a plane, our problem can be
defined as minimally covering a polygon i.e. the plane, using other
polygons of arbitrary geometry i.e. our image projections, with the
added constraint of minimizing our cost function between chosen
images.  This is a complex problem, though we can take a number of
steps to simplify it. Given that our wall-texture candidate images are
all taken from a head-on angle, and assuming only minor rotations are
made during localization refinement, we can reason that their
projections onto the plane are approximately rectangular, as in Figure
\ref{fig:highCameraAngle}. By discarding the minor excess texture and
cropping them all to be rectangular, our problem becomes the
conceptually simpler one of filling a polygon with rectangles, such
that the sum of all edge costs between each pair of rectangles is
minimal. We thus also retain the advantages of working with
rectangular units, as explained in Section
\ref{sec:simpleTextureMapping}.

The location and orientation of the cameras on our backpack is such
that our images nearly always contain the entirety of the floor to
ceiling range of wall planes. Images are therefore rarely projected
with one above the other when texturing wall planes, which correspond
to the optimal planes we are working with.  In essence, we need only
to ensure horizontal coverage of our planes, as our images provide
full vertical coverage themselves. We can thus construct a Directed
Acyclic Graph (DAG) from the images, with edge costs defined by our
cost function above, and solve a simple shortest path problem to find
an optimal subset of images with regard to the cost functions.

@@

We now solve the shortest path problem from the start node to the end
node \cite{dijkstra}. This provides a set of images completely
covering the plane horizontally, while minimizing the cost of the
seams between images.

>>>>

%%%%%%%%% ABSTRACT
\begin{abstract}
  Automated 3D modeling of building interiors is useful in
  applications such as virtual reality and environment
  mapping. Applying textures to these models is an important step in
  generating photorealistic visualizations of data collected by
  modeling systems.  Camera pose recovery in such systems often
  suffers from inaccuracies, resulting in visible discontinuties when
  different images are projected adjacently onto a plane for
  texturing. We propose two approaches for reducing discontinuities in
  texture mapping 3d models made of planar surfaces. The first one is
  tile based and can be used where camera axes are at arbitrary angles
  with respect to the plane normals of the surfaces. The second one
  results in a more seamless texture but is only applicable where
  camera axes for imagery are closely aligned with plane normals. The
  effectiveness of our approaches are demonstrated on two indoor
  datasets.
\end{abstract}

@@

The first step in automated 3D modeling is the physical scanning of
the environment's geometry. An indoor modeling system must be able to
recover camera poses within an environment while simulatenously
reconstructing the 3D structure of the environment itself
\cite{chen2010indoor, hz, kua2012loopclosure, liu2010indoor}.This is
known as the simultaneous localization and mapping (SLAM) problem, and
is generally solved by taking readings from laser range scanners,
cameras, and inertial measurement units (IMUs) at multiple locations
within the environment.

Mounting such devices on a human-carried platform provides unique
advantages over vehicular-based systems on wheels in terms of agility and
portability, but can also result in much
larger localization error. As a result, common methods for texture
mapping generally produce poor results.

In this paper, we present a number of approaches to texture mapping 3D models of indoor environments made of planar surfaces in the presence of uncertainty and noise in camera poses. In particular, we consider a human-operated backpack system with a number of laser range scanners as well as 2 cameras facing left and right, each equipped with fisheye lenses reaching an approximately 180$^{\circ}$ field of view. Considering that the human operator walks at about 1 meter/second, and images are captured at 5 Hz, there is a considerable amount of overlap between successive images.

@@

Using data collected by the onboard sensors and applying multiple localization
and loop-closure algorithms \cite{chen2010indoor, kua2012loopclosure,
  liu2010indoor}, the backpack is then localized over its
data collection period. This requires recovering the 6 degrees of freedom for the backpack as well as all its sensors, including the cameras mounted on it. \footnote{In this paper, we use the terms localization and pose recovery interchangeably, in that they both refer to recovering position and orientation.} Once this is complete, the data from the laser range scanners is used to generate a 3D point cloud of the surrounding environment. 
Approximate normal vectors for each point in the
point cloud are then calculated by gathering neighboring points within
a small radius and processing them through principal component
analysis. These normal vectors allow for the classification and
grouping of adjacent points into structures such as walls, ceilings,
floors, and staircases. A RANSAC algorithm is then employed to fit
polygonal planes to these structured groupings of points, resulting in
a fully planar model \cite{sanchez2012point}. This model, consisting
of multiple 2D polygonal planes in 3D space, along with the set of
images captured by the backpack's cameras and their noisy 3D poses,
can be considered the input to our texture mapping problem.

The remainder of the paper is organized as follows. Section
\ref{sec:simpleTextureMapping} describes the general problem of
texture mapping and examines two variants of a tile-based mapping approach and their
shortcomings.  Section \ref{sec:existingApproaches} demonstrates two
previous attempts at improving texture alignment, and demonstrates their
inadequacies for our datasets.  Section \ref{sec:proposedApproach}
presents our proposed approach to texture mapping, combining an
improved localization refinement process with two image selection
approaches. Section \ref{sec:resultsAndConclusions} contains results
and conclusions.

@@

\section{Tile-Based Texture Mapping}
\label{sec:simpleTextureMapping}
\begin{figure}
  \centering
  \includegraphics[height=2in]{Projection.pdf}
  \caption{Planes are specified in 3D space by four corners $C_1$ to
    $C_4$. Images are related to each plane through the camera
    matrices $P_{1..M}$. }
  \label{fig:projection}
\end{figure}

The geometry of the texture mapping process for a plane is shown in
Figure \ref{fig:projection}.  As described earlier, we are provided
with a set of $M$ images with which we must texture our target
plane. Each image has a camera matrix $P_i$ for $i=1..M$, which
translates a 3D point in the world coordinate system to a 2D point or
pixel in image $i$'s coordinates. A camera matrix $P_i$ is composed of
the camera's intrinsic parameters, such as focal length and image
center, as well as extrinsic parameters which specify the rotation and
translation of the camera's position in 3D world coordinates at the
time that image $i$ is taken. These extrinsic parameters are
determined by the backpack hardware and localization algorithms
\cite{chen2010indoor, liu2010indoor, kua2012loopclosure} and are quite
noisy.

Our goal is to texture each plane using images captured by the
backpack system, while eliminating any visual discontinuities or seams
that would suggest that the plane's texture is not composed of a
single continuous image.

@@

In order to select an image for texturing tile $t$, we must first
gather a list of candidate images that contain all four of its
corners, which we can quickly check by projecting $t$ into each image
using the projection method above. Furthermore, each candidate image
must have been taken at a time when its camera had a clear
line-of-sight to $t$, which can be calculated using standard
ray-polygon intersection tests between the camera location, the center
of $t$, and every other plane, all in world coordinates
\cite{rayintersection}.

Once we have a list of candidate images for $t$, we must define a
scoring function in order to compare images and objectively select the
best one. Since camera localization errors compound over distance, we
wish to minimize the distance between cameras and the plane they
texture. Additionally, we desire images that are projected
perpendicularly onto the plane, maximizing the resolution and amount
of useful texture available in their projections. In other words, we
wish to minimize the angle between the plane normal and the camera
axis for images selected for texture mapping. These two criteria can
be met by maximizing the function $\frac{1}{d} (-1 \cdot \vec{C})
\cdot \vec{N}$ with the parameters shown in Figure
\ref{fig:scoringFunction}. Specifically, $d$ is the distance between
the centers of a camera and a tile, and $\vec{N}$ and $\vec{C}$ are
the directions of the plane's normal and the camera axis respectively.

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{scoringFunction.pdf}
  \caption{We wish to minimize camera angle $\alpha$ and distance $d$}
  \label{fig:scoringFunction}
\end{figure}

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{wall1_naive.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=3in]{wall1_cache_full.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=3in]{wall1_cache_full_shifted.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=3in]{wall1_dynprog_noblend.jpg}}
  \caption{(a): Direct mapping. (b): Mapping with caching. (c):
    Mapping with caching after image alignment. (d): Seam minimization approach (including image alignment)}
  \label{fig:compareAll}
\end{figure}

As Figure \ref{fig:compareAll}(a) demonstrates, this approach leads
to the best texture for each tile independently, but overall results
in many image boundaries with abrupt discontinuities, due to
significant misalignment between images, as a result of camera pose
inaccuracies.

\subsection{Mapping with Caching}
\label{sec:mappingWithCaching}
Since discontinuities occur where adjacent tiles select non-aligned
images that do not align, it makes sense to take into account image
selections made by neighboring tiles while selecting the best image
for a given tile. By using the same image across tile boundaries, we
can eliminate a discontinuity altogether. If this is not possible
because a tile is not visible in images chosen by its neighbors, using
similar images is likely to result in less noticeable discontinuities.

Similar to a caching mechanism, we select the best image for a tile
$t$ by searching through two subsets of images for a good candidate,
before searching through the entire set. The first subset of images is
those selected by adjacent tiles that have already been textured. We
must first check which images can map to $t$, and then of those, we
make a choice according to the same scoring function. Before reusing
this image, we ensure it meets a threshold, which we set
unconditionally to $\alpha < 45^\circ$, to be considered a viable
image, with $\alpha$ as the camera angle as shown in Figure
\ref{fig:scoringFunction}.

If no satisfactory image is found in the first subset, we then check
our second subset of images, which consists of images that were taken
near the images in the first subset, both spatially and
temporally. These images are not the same as the ones used for
neighboring tiles, but they were taken at a similar location and time,
suggesting that their localization and projection are very
similar. Again, if no good image is found according to the same
threshold, we must then search the entire set of candidate images,
selecting based on the criteria from Figure \ref{fig:scoringFunction}.

The results of this caching approach are shown in Figure
\ref{fig:compareAll}(b). As compared to Figure
\ref{fig:compareAll}(a), discontinuities are reduced overall, but
the amount of remaining seams suggests that image selection alone
cannot produce seamless textures. Camera matrices, or the image
projections themselves have to be adjusted in order to reliably
generate clean textures.

\section{Existing Approaches to Image-Aligned Texture Mapping}
\label{sec:existingApproaches}
In order to produce seamless texture mapping, either camera matrices
need to be refined such that their localization is pixel accurate, or
image stitching techniques need to be applied to provide this
illusion.

Before examining these approaches, we first obtain a set of images to
work with.  Rather than perform camera or image adjustments across the
many thousands of images acquired in a typical data collection, we opt
to work with the more limited set of images corresponding to those
chosen by the direct mapping approach, without caching (Section
\ref{sec:directMapping}. This set of images constitutes a good
candidate set for generating a final seamless texture since it meets
three important criteria. First, each tile on our planeis covered by
at least one image in this set; this ensures no holes in our final
texture. Second, images are all selected according to the scoring
function in Figure \ref{fig:scoringFunction}, ensuring good camera
distances and angles. Third, as a side result of the scoring function,
selected images are only viable candidates for the tiles near their
center of projection. Thus, there should be plenty of overlap between
selected images, allowing for some degree of shifting without
resulting in holes, as well as area for blending between them. With
this set of images, we now review two existing approaches towards
refining and combining their projections.

\subsection{Image Mosaicing}
\label{sec:imageMosaicing}

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{panoMy.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=3in]{Graph_crop.pdf}}
  \caption{Both image mosaicing (a) and the graph-based localization refinement algorithm from
    \cite{chen2010indoor} (b) suffer from the problem of compounding
    errors.}
  \label{fig:mosaic3D}
\end{figure}

@@

Figure \ref{fig:mosaic3D}(a) shows the output of the AutoStitch software
package which does homography-based image mosaicing. This plane is
nearly a best-case scenerio with many features spread uniformly across
it. Even so, the mosaicing produces errors that causes straight lines
to appear as waves on the plane, despite the fact that it was
generated after careful hand tuning. Many planes with fewer features
simply failed outright. Thus, image mosaicing is not a robust enough
solution for reliably texture mapping our dataset.

\subsection{Image-Based 3D Localization Refinement}
\label{sec:imageBased3DRefinement}

Another approach is to refine the camera matrices using image
correspondences to guide the process. Each image's camera matrix has 6
degrees of freedom that can be adjusted. Previous work on this problem
attempted to refine camera matrices by solving a non-linear
optimization problem \cite{liu2010indoor}. This process is specific to
the backpack system which generated our dataset, as it must be run
during backpack localization\cite{liu2010indoor,
  chen2010indoor}. This approach suffers from a similar
error propagation problem as shown in Figure \ref{fig:mosaic3D}(b).

\begin{figure}
  \centering
  \includegraphics[width=3in]{pipeline.pdf}
  \caption{Our proposed method for seamless texture mapping.}
  \label{fig:pipeline}
\end{figure}

\section{Proposed Method for Seamless Texture Mapping}
\label{sec:proposedApproach}
In this section, we will describe our proposed method for seamless
texture mapping. Our approach consists of 4 steps, as seen in Figure
\ref{fig:pipeline}.  First, we choose a subset of $M$ images to work
with via the direct mapping approach of Section
\ref{sec:directMapping}. Next, we perform image rotation and shifting
in order to maximize SIFT feature matches in overlapping areas between
images. We then project these images onto our plane, applying textures
either with the tile caching method from Section
\ref{sec:mappingWithCaching}, or with the more specialized method to
be described in Section \ref{sec:seamMinimization}. This choice
depends on the availability of head-on images for a given plane. We
then finish by applying linear alphar blending to smooth out any
remaining seams.

\subsection{Image Projection and Rotation}
\label{sec:projectionAndRotation}
We begin with the projection of all images onto the plane. This is
done in the same way as the approaches in Section
\ref{sec:simpleTextureMapping}. We then perform rotations on these
projections, as adjacent images with different orientations result in
strong discontinuities.

These rotations are accomplished by using Hough transforms, which
detect the presence and orientation of linear features in our
images. Rather than match the orientation of such features in each
image, we simply apply rotations such that the strongest near-vertical
features are made completely vertical. This is effective for vertical
planes in indoor models, since they usually consist of parallel
vertical lines corresponding to doors, wall panels, rectangular
frames, etc. If features in the environment are not vertical, or are
not parallel to eachother, this step is skipped.

\subsection{Image Pose Refinement}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images using
SIFT feature matching \cite{lowe1999object}. The SIFT matches
determine $dx$ and $dy$ distances between each pair of features for
two images on the plane, though these distances may not always be the
same for different features.

@@

@@

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and match each other
with far greater accuracy. Applying the tile caching method from
Section \ref{sec:mappingWithCaching} on these re-localized images
results in the significant improvements shown in Figure
\ref{fig:compareAll}(c), as compared to Figure \ref{fig:compareAll}(d).

@@

\begin{figure}
  \centering
  \includegraphics[width=3in]{highCameraAngle.pdf}
  \caption{(1) projects perpendicularly onto a wall and provides
    uniformly high resolution textures over its entire projection
    (2). (3) on the other hand projects at a poor angle onto the
    ceiling (4), resulting in usable texture at point A, but
    increasingly low resolution texture towards point B.}
  \label{fig:highCameraAngle}
\end{figure}

As mentioned earlier, our backpack system has 2 cameras with
$180^\circ$ fisheye lenses facing to the right and left. Since the
backpack operator only takes care to fully scan each wall and not
necessarily the entirety of each ceiling and floor, cameras at higher
angles with respect to plane normal vectors must be used for texturing
large areas of floor and ceiling planes. These high camera angles
translate into images that span extremely large areas once
projected. Such image projections have good scores (from Figure
\ref{fig:scoringFunction}) for certain areas on the plane, but much
worse scores for other areas. This can be seen in Figure
\ref{fig:highCameraAngle}. The tiling approach used in Section
\ref{sec:simpleTextureMapping} was thus a good choice for texturing,
as it allowed us to only use the parts of image projections that were
good choices for their respective plane locations, e.g. areas near
point A in Figure \ref{fig:highCameraAngle}, but not near point B.

For wall planes however, we have images all taken from close distances
and more head-on angles, and thus higher resolution, near-rectangular
projections to work with. As a result, there is less deviation over
the score of each tile within an image, as well as over the average
scores of all images. This means that the scoring criteria from Figure
\ref{fig:scoringFunction} is less relevant to walls, which enjoy an
abundance of head-on images.

@@

\subsubsection{Occlusion Masking}
\label{sec:occlusionMasking}
To begin with, we need to ensure that our images contain only content
that should be mapped onto the target plane in question. The tiling
approach used previously only checks occlusion for each tile as it is
being textured. For our new approach, we prefer entire images, so we
need to perform occlusion checks over the entirety of each image to
determine available areas for texture mapping.

@@

\subsubsection{Image Selection}
\label{sec:imageSelection}
To determine the set of images that results in the cleanest texture,
we need a cost function to evaluate the visibiilty of seams between
images in our set. A straightforward cost function that accomplishes
this is the sum of squared pixel differences in overlapping regions
between all pairs of images, after they have been aligned as described
in Section \ref{sec:projectionAndRotation}. Minimizing this cost
function encourages image boundaries to occur either in featureless
areas, such as bare walls, or in areas where images match extremely
well.

With the cost function defined, we must now select the set of images
for which the overall cost function is minimized. Since nearly all the
images for these optimal cases are head on, the best strategy to
minimize seams is to choose as few images as possible while texturing
a given plane. Thus,to cover the entirety of a plane, our problem can
be defined as minimally covering a polygon i.e. the plane, using other
polygons of arbitrary geometry i.e. our image projections, with the
added constraint of minimizing our cost function between chosen
images.  This is a complex problem, though we can take a number of
steps to simplify it. Given that our wall-texture candidate images are
all taken from a head-on angle, and assuming only minor rotations are
made during localization refinement, we can reason that their
projections onto the plane are approximately rectangular, as in Figure
\ref{fig:highCameraAngle}. By cropping them all to be rectangular, our
problem becomes the conceptually simpler one of filling a polygon with
rectangles, such that the sum of all edge costs between each pair of
rectangles is minimal. We thus also retain the advantages of working
with rectangular units, as explained in Section
\ref{sec:simpleTextureMapping}.

The location and orientation of the cameras on our backpack is such
that our images nearly always contain the entirety of the floor to
ceiling range of wall planes. Images are therefore rarely projected
with one above another when texturing wall planes, which correspond to
the optimal case we are working with.  In essence, we need only to
ensure horizontal coverage of our planes, as our images provide full
vertical coverage themselves. We can thus construct a Directed Acyclic
Graph (DAG) from the images, with edge costs defined by our cost
function above, and solve a simple shortest path problem to find an
optimal subset of images with regard to the cost functions
\cite{dijkstra}.

@@

We now solve the shortest path problem from the start node to the end
node. This provides a set of images completely covering the plane
horizontally, while minimizing the cost of seams between images.

In rare cases where the vertical dimension of the plane is not
entirely covered by one or more chosen images, we are left with holes
where no images are selected to texture. To address this, we simply
modify the edges chosen by our shortest path solution such that they
have higher cost than all other edges in our DAG, and solve for a new
shortest path. This ensures that these edges will not be chosen again
unless they are the only available option. Our second shortest path
solution will result in a new set of chosen images, of which we only
retain those that cover areas not covered by the first set. This
process can be repeated as many times as needed, until holes are no
longer present. This method is not as optimal as a 2D-coverage
solution would be, but it is a quick approximation, and adequately
handles the few holes we run into.
